```{r , warning=FALSE, message=FALSE, echo=FALSE}
library(data.table)

library(sandwich)
library(lmtest)

library(ggplot2)
library(patchwork)

library(foreign)

library(knitr)
library(tidyverse)
library(kableExtra)
library(stargazer)
inline_reference <- "r inline_reference"
```

# Analysis

## Intent to Treat (ITT) Effect
Our first step in evaluating the outcomes of our experiments is to estimate the Intent to Treat ($ITT$). The $ITT$ measures the outcomes of our experiment based on treatment assignment, without considering compliance rate. The $ITT$ is defined as:

\begin{quote}
  \textit{$ITT$ = $E[Y_i(z = 1)] - E[Y_i(z = 0)]$}
\end{quote}

where z signifies the assignment of the subject. Below shows the ITT across six of our ten 10 different outcome variables for each experiment. The table containing all ten outcomes can be found in the Appendix.

```{r, include=FALSE}
exp_data      <- fread('../data/exp_data.csv')
exp_data_main <- fread('../data/exp_data_gpt.csv')
exp_data_main <- exp_data_main[maia_name %in% c('chess4gerry', 'ucb_123', 'sandman353', 'bcu_555')]

```

```{r feature engineer, include=FALSE}
# one-hot encode maia win
exp_data[ , maia_win := ifelse(winner == 'maia', 1, 0)]
exp_data_main[ , maia_win := ifelse(winner == 'maia', 1, 0)]

# one-hot encode maia checkmate
exp_data[ , maia_checkmate := ifelse(winner == 'maia' & status == 'mate', 1, 0)]
exp_data_main[ , maia_checkmate := ifelse(winner == 'maia' & status == 'mate', 1, 0)]

# one-hot encode opp checkmate
exp_data[ , opp_checkmate := ifelse(winner == 'opp' & status == 'mate', 1, 0)]
exp_data_main[ , opp_checkmate := ifelse(winner == 'opp' & status == 'mate', 1, 0)]

# one-hot encode opp resign
exp_data[ , opp_resign := ifelse(winner == 'maia' & status == 'resign', 1, 0)]
exp_data_main[ , opp_resign := ifelse(winner == 'maia' & status == 'resign', 1, 0)]

# one-hot encode draw
exp_data[ , draw := ifelse(status == 'draw', 1, 0)]
exp_data_main[ , draw := ifelse(status == 'draw', 1, 0)]

# one-hot encode treatment
exp_data[ , treat := ifelse(chat_type == 'Chatty', 1, 0)]
exp_data_main[ , treat := ifelse(chat_type == 'ChatGPT', 1, 0)]

```


```{r define measures, include=FALSE}
# define the outcome metrics for stargazers
measures <- c("Maia Win",
              #"Maia Checkmate",
              #"Opp Checkmate",
              "Opp Resigns",
              #"Draw",
              "Opp Acc",
              "Opp Blunders",
              "Opp Mistakes",
              #"Opp ACPL",
              "Opp Avg Mv Time")
```


```{r ITT calc exp1 TvC, echo=FALSE,include=FALSE}
exp_data_c_t <- exp_data[chat_type %in% c("Chatty", "Control") , ]

# only include rows where respond != NA
exp_data_c_t <- exp_data_c_t[complete.cases(exp_data_c_t$respond), ]

itt_prelim_win           <- exp_data_c_t[ , lm(maia_win ~ treat)]
itt_prelim_maia_mate     <- exp_data_c_t[ , lm(maia_checkmate ~ treat)]
itt_prelim_opp_mate      <- exp_data_c_t[ , lm(opp_checkmate ~ treat)]
itt_prelim_opp_resign    <- exp_data_c_t[ , lm(opp_resign ~ treat)]
itt_prelim_draw          <- exp_data_c_t[ , lm(draw ~ treat)]
itt_prelim_opp_acc       <- exp_data_c_t[ , lm(opp_acc ~ treat)]
itt_prelim_opp_blunders  <- exp_data_c_t[ , lm(opp_blunders ~ treat)]
itt_prelim_opp_mistakes  <- exp_data_c_t[ , lm(opp_mistakes ~ treat)]
itt_prelim_opp_acpl      <- exp_data_c_t[ , lm(opp_acpl ~ treat)]
itt_prelim_opp_move_time <- exp_data_c_t[ , lm(opp_avg_move_time ~ treat)]

# add robust standard errors

itt_prelim_win_rse           <- coeftest(itt_prelim_win, vcov = vcovHC)
itt_prelim_maia_mate_rse     <- coeftest(itt_prelim_maia_mate, vcov = vcovHC)
itt_prelim_opp_mate_rse      <- coeftest(itt_prelim_opp_mate, vcov = vcovHC)
itt_prelim_opp_resign_rse    <- coeftest(itt_prelim_opp_resign, vcov = vcovHC)
itt_prelim_draw_rse          <- coeftest(itt_prelim_draw, vcov = vcovHC)
itt_prelim_opp_acc_rse       <- coeftest(itt_prelim_opp_acc, vcov = vcovHC)
itt_prelim_opp_blunders_rse  <- coeftest(itt_prelim_opp_blunders, vcov = vcovHC)
itt_prelim_opp_mistakes_rse  <- coeftest(itt_prelim_opp_mistakes, vcov = vcovHC)
itt_prelim_opp_acpl_rse      <- coeftest(itt_prelim_opp_acpl, vcov = vcovHC)
itt_prelim_opp_move_time_rse <- coeftest(itt_prelim_opp_move_time, vcov = vcovHC)

itt_prelim_models <- list(itt_prelim_win_rse,
                       itt_prelim_opp_resign_rse,
                       itt_prelim_opp_acc_rse,
                       itt_prelim_opp_blunders_rse,
                       itt_prelim_opp_mistakes_rse,
                       itt_prelim_opp_move_time_rse)

```

```{r, results='asis', echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, font.size = "small"}
stargazer(itt_prelim_models,
          type = 'latex',
          header = FALSE,          
          dep.var.labels = 'Outcome Variables',
          column.labels = measures,
          covariate.labels = "Treatment")
```


```{r, echo=FALSE,include=FALSE}
exp_data_main[chat_type %in% c("ChatGPT", "Control GPT") , ]

# only include rows where respond != NA
exp_data_main <- exp_data_main[complete.cases(exp_data_main$respond), ]

itt_main_win           <- exp_data_main[ , lm(maia_win ~ treat)]
itt_main_maia_mate     <- exp_data_main[ , lm(maia_checkmate ~ treat)]
itt_main_opp_mate      <- exp_data_main[ , lm(opp_checkmate ~ treat)]
itt_main_opp_resign    <- exp_data_main[ , lm(opp_resign ~ treat)]
itt_main_draw          <- exp_data_main[ , lm(draw ~ treat)]
itt_main_opp_acc       <- exp_data_main[ , lm(opp_acc ~ treat)]
itt_main_opp_blunders  <- exp_data_main[ , lm(opp_blunders ~ treat)]
itt_main_opp_mistakes  <- exp_data_main[ , lm(opp_mistakes ~ treat)]
itt_main_opp_acpl      <- exp_data_main[ , lm(opp_acpl ~ treat)]
itt_main_opp_move_time <- exp_data_main[ , lm(opp_avg_move_time ~ treat)]

# add robust standard errors

itt_main_win_rse           <- coeftest(itt_main_win, vcov = vcovHC)
itt_main_maia_mate_rse     <- coeftest(itt_main_maia_mate, vcov = vcovHC)
itt_main_opp_mate_rse      <- coeftest(itt_main_opp_mate, vcov = vcovHC)
itt_main_opp_resign_rse    <- coeftest(itt_main_opp_resign, vcov = vcovHC)
itt_main_draw_rse          <- coeftest(itt_main_draw, vcov = vcovHC)
itt_main_opp_acc_rse       <- coeftest(itt_main_opp_acc, vcov = vcovHC)
itt_main_opp_blunders_rse  <- coeftest(itt_main_opp_blunders, vcov = vcovHC)
itt_main_opp_mistakes_rse  <- coeftest(itt_main_opp_mistakes, vcov = vcovHC)
itt_main_opp_acpl_rse      <- coeftest(itt_main_opp_acpl, vcov = vcovHC)
itt_main_opp_move_time_rse <- coeftest(itt_main_opp_move_time, vcov = vcovHC)

itt_main_models <- list(itt_main_win_rse,
                        itt_main_opp_resign_rse,
                        itt_main_opp_acc_rse,
                        itt_main_opp_blunders_rse,
                        itt_main_opp_mistakes_rse,
                        itt_main_opp_move_time_rse)

```

```{r, results='asis', echo=FALSE, warning=FALSE, message=FALSE, include=TRUE}
stargazer(itt_main_models,
          type = 'latex',
          header = FALSE,
          dep.var.labels = 'Outcome Variables',
          column.labels = measures,
          covariate.labels = "Treatment")

```

## Non-compliance
In our experiment, we only tested for one-sided compliance. In particular, we tested whether someone was intended to receive treatment, but did not engage in the chat room. Conversely, our bot was not programmed to engage with opponents if they chatted first, so we did not have concerns of non-compliance where someone may have been targeted for control, but instead received treatment.


## Complier Average Causal Effect (CACE)
While the $ITT$ allows us to get a glimpse into potential treatment effects, it does not consider compliers and non-compliers. As seen below, not everyone received their targeted dose of treatment, which we now take into account in calculating the Complier Average Causal Effect ($CACE$).


```{r, include = FALSE}
DT1 = data.table(
  response_rate = c(4.5, 35.5, 30.1),
  treatment = c("Control", "Placebo", "Treatment")
)
p_exp_graph <- ggplot(DT1, aes(x= treatment, y = response_rate, label = response_rate)) +
  geom_col(fill="royalblue3") +
  ylab("Response rate %") +
  xlab("") +
  ggtitle("Preliminary Experiment") +
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 7)) +
  ylim(0,40)
DT2 = data.table(
  response_rate = c(.45, 33.2),
  treatment = c("Control", "ChatGPT")
)
m_exp_graph <- ggplot(DT2, aes(x= treatment, y = response_rate, label = response_rate)) +
  geom_col(fill="tomato2") +
  ylab("Response rate %") +
  xlab("") +
  ggtitle("Main Experiment") +
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 7)) +
  ylim(0,40)
```

```{r, include=TRUE, echo=FALSE, fig.height=3.5, fig.cap= NULL, fig.pos='!b',fig.cap="\\label{fig:figs}Response Rates by Experiment"}
p_exp_graph | m_exp_graph
```

The $CACE$ is calculated using the following formula:

\begin{quote}
  \textit{$CACE$ = $ITT$/$ITT_d$}
\end{quote}

where $ITT_d$ is the proportion of compliers in the treatment group (the “take-up rate”). In our case, we can see above that the take-up rates are approximately 30% and 33% for the preliminary and main experiments, respectively.

```{r cace function, echo=FALSE, include=FALSE}
# define function to calculate cace
get_cace <- function(data, treat_var, compliance_var, outcome_var) {
  # first stage regression
  first_stage <- lm(paste(compliance_var, "~", treat_var), data = data)
  
  # extract predicted values of compliance from the first stage
  predicted_compliance <- predict(first_stage)
  
  # create a new data frame with the predicted compliance values
  data_with_predicted <- cbind(data, predicted_compliance)
  
  # Second stage regression
  cace_model <- lm(paste(outcome_var, "~ predicted_compliance"), data = data_with_predicted)
  
  # Obtain summary statistics for the CACE model
  cace_model
}

```

```{r cace calc exp1}
cace_prelim_win               <- get_cace(exp_data_c_t, "treat", "respond", "maia_win")
cace_prelim_maia_mate         <- get_cace(exp_data_c_t, "treat", "respond", "maia_checkmate")
cace_prelim_opp_mate          <- get_cace(exp_data_c_t, "treat", "respond", "opp_checkmate")
cace_prelim_opp_resign        <- get_cace(exp_data_c_t, "treat", "respond", "opp_resign")
cace_prelim_draw              <- get_cace(exp_data_c_t, "treat", "respond", "draw")
cace_prelim_opp_acc           <- get_cace(exp_data_c_t, "treat", "respond", "opp_acc")
cace_prelim_opp_blunders      <- get_cace(exp_data_c_t, "treat", "respond", "opp_blunders")
cace_prelim_opp_mistakes      <- get_cace(exp_data_c_t, "treat", "respond", "opp_mistakes")
cace_prelim_opp_acpl          <- get_cace(exp_data_c_t, "treat", "respond", "opp_acpl")
cace_prelim_opp_move_time     <- get_cace(exp_data_c_t, "treat", "respond", "opp_avg_move_time")

# add robust standard errors

cace_prelim_win_rse           <- coeftest(cace_prelim_win, vcov = vcovHC)
cace_prelim_maia_mate_rse     <- coeftest(cace_prelim_maia_mate, vcov = vcovHC)
cace_prelim_opp_mate_rse      <- coeftest(cace_prelim_opp_mate, vcov = vcovHC)
cace_prelim_opp_resign_rse    <- coeftest(cace_prelim_opp_resign, vcov = vcovHC)
cace_prelim_draw_rse          <- coeftest(cace_prelim_draw, vcov = vcovHC)
cace_prelim_opp_acc_rse       <- coeftest(cace_prelim_opp_acc, vcov = vcovHC)
cace_prelim_opp_blunders_rse  <- coeftest(cace_prelim_opp_blunders, vcov = vcovHC)
cace_prelim_opp_mistakes_rse  <- coeftest(cace_prelim_opp_mistakes, vcov = vcovHC)
cace_prelim_opp_acpl_rse      <- coeftest(cace_prelim_opp_acpl, vcov = vcovHC)
cace_prelim_opp_move_time_rse <- coeftest(cace_prelim_opp_move_time, vcov = vcovHC)

cace_prelim_models <- list(cace_prelim_win_rse,
                    cace_prelim_opp_resign_rse,
                    cace_prelim_opp_acc_rse,
                    cace_prelim_opp_blunders_rse,
                    cace_prelim_opp_mistakes_rse,
                    cace_prelim_opp_move_time_rse
                    )
```

```{r, include=TRUE, results='asis', echo=FALSE, warning=FALSE}

stargazer(cace_prelim_models,
          dep.var.labels = 'Outcome Variables',
          column.labels = measures,
          covariate.labels = "Treatment",
          type = 'latex',
          header = FALSE)
     
```

```{r cace calc exp2}
# remove NAs

cace_main_win               <- get_cace(exp_data_main, "treat", "respond", "maia_win")
cace_main_maia_mate         <- get_cace(exp_data_main, "treat", "respond", "maia_checkmate")
cace_main_opp_mate          <- get_cace(exp_data_main, "treat", "respond", "opp_checkmate")
cace_main_opp_resign        <- get_cace(exp_data_main, "treat", "respond", "opp_resign")
cace_main_draw              <- get_cace(exp_data_main, "treat", "respond", "draw")
cace_main_opp_acc           <- get_cace(exp_data_main, "treat", "respond", "opp_acc")
cace_main_opp_blunders      <- get_cace(exp_data_main, "treat", "respond", "opp_blunders")
cace_main_opp_mistakes      <- get_cace(exp_data_main, "treat", "respond", "opp_mistakes")
cace_main_opp_acpl          <- get_cace(exp_data_main, "treat", "respond", "opp_acpl")
cace_main_opp_move_time     <- get_cace(exp_data_main, "treat", "respond", "opp_avg_move_time")

# add robust standard errors

cace_main_win_rse           <- coeftest(cace_main_win, vcov = vcovHC)
cace_main_maia_mate_rse     <- coeftest(cace_main_maia_mate, vcov = vcovHC)
cace_main_opp_mate_rse      <- coeftest(cace_main_opp_mate, vcov = vcovHC)
cace_main_opp_resign_rse    <- coeftest(cace_main_opp_resign, vcov = vcovHC)
cace_main_draw_rse          <- coeftest(cace_main_draw, vcov = vcovHC)
cace_main_opp_acc_rse       <- coeftest(cace_main_opp_acc, vcov = vcovHC)
cace_main_opp_blunders_rse  <- coeftest(cace_main_opp_blunders, vcov = vcovHC)
cace_main_opp_mistakes_rse  <- coeftest(cace_main_opp_mistakes, vcov = vcovHC)
cace_main_opp_acpl_rse      <- coeftest(cace_main_opp_acpl, vcov = vcovHC)
cace_main_opp_move_time_rse <- coeftest(cace_main_opp_move_time, vcov = vcovHC)

cace_main_models <- list(cace_main_win_rse,
                    cace_main_opp_resign_rse,
                    cace_main_opp_acc_rse,
                    cace_main_opp_blunders_rse,
                    cace_main_opp_mistakes_rse,
                    cace_main_opp_move_time_rse
                    )
```

```{r, include=TRUE, results='asis', echo=FALSE, warning=FALSE}

stargazer(cace_main_models,
          dep.var.labels = 'Outcome Variables',
          column.labels = measures,
          covariate.labels = "Treatment",
          type = 'latex',
          header = FALSE)
     
```

## Benjamini-Hochberg (BH) Procedure
While some test results in the tables above show statistical significance at the 0.1 $p$ value threshold, we need to adjust due to the number of tests that we conducted. Since we performed ten statistical tests, we have an increased chance of finding a $p$ value less than or equal to 0.1 by chance, compared to running one or two tests. The Benjamini-Hochberg (BH) procedure is one way to try to control for the false discovery rate (FDR) between all of our tests. Unlike the False Positive Rate, which is used in the Bonferroni method, the FDR is the expected proportion of false positives among *all* positives which rejected the null hypothesis

In this method, we take the $p$ values from each of our $CACE$ outcomes and compare them to their associated $BH_{critical}$ value. The $BH_{critical}$ value is defined as:

\begin{quote}
  \textit{$BH_{critical}$ = ($i$/$m$) $*$ $Q$}
\end{quote}

where $i$ is the rank of the $p$ value (i.e., the smallest value has $i=1$, etc.), $m$ is is the number of $p$ values (i.e., ten), and $Q$ is the allowable FDR (i.e., 0.05). In this test, if a $p$ value is less than its $BH_{critical}$ value, then the test is significant. Otherwise, the $p$ values that we found are likely the result of a false discovery.

In both our Preliminary Experiment and Main Experiment, there were zero $p$ values less than their associated $BH_{critical}$ value, and thus we fail to reject the null hypothesis under both experiment conditions.

