---
title: "Power Analysis"
author: "Jared Feldman, Jerry Gonzalez, Reed Evans, Randy Louie"
date: "10/10/2023"
output:
  pdf_document:
    toc: yes
  bookdown::pdf_document2:
    toc: yes
    number_sections: yes
subtitle: Datasci 241
fig.caption: yes
keep_tex: yes
header-includes: \usepackage{float}
---

\newpage
\setcounter{page}{1}

```{r package loads, warning=FALSE, message=FALSE, include=FALSE}
library(data.table)
library(sandwich)
library(lmtest)
library(ggplot2)
library(knitr)
library(tidyverse)
```
## Justification for simulation variables
We performed a pilot test of 10 games with our chess bot that plays at a consistent 1500 (average) level against real opponents on lichess.com.  We found that the opponents played at an average of 72% accuracy and had a 12.3 standard deviation in their accuracy rating.  We used these values in our power analysis.  

We are using a conservative estimate of 5-15% effect size for our power analysis.  We found in the paper *Trash-talking: Competitive incivility motivates rivalry, performance, and unethical behavior* by Jeremy A. Yip, Maurice E. Schweitzer, and Samir Nurohamed.  https://www.sciencedirect.com/science/article/pii/S0749597816301157 

They found that trash talking had an an indirect effect on competitive performance through creation of a perceived rivalry between the players. On page 131, they show that there was an effect of b=.32, with a 95% confidence interval of 0.02, 0.87. Trash talking resulted in the opponent performing better.  

## Create simulation function using linear regression
```{r Function-for-simulation, include = TRUE}
simulate_regression <- function(recruits, mean_bad_moves, effect_size, sd_bad_moves){
    d <- data.table(
      n = 1:recruits)
    
    half_recruits <- floor(recruits/2)
    
    ## no treatment data 
    d_1 <- data.table(
      id    = 1:half_recruits, 
      treat = 0)
    
    ## assign num bad moves to treatment group
    d_1[ , Bad_moves := rnorm(.N, mean=mean_bad_moves, sd=sd_bad_moves)]
    
    ## treatment data 
    d_2 <- data.table(
      id    = (half_recruits+1):recruits, 
      treat = 1)
    
    ## assign num bad moves to treatment group
    d_2[ , Bad_moves := rnorm(.N, mean=mean_bad_moves*(1+effect_size), sd=sd_bad_moves)]
    
    ## Stack data frames 
    d <- rbind(d_1, d_2)
    
    model_1    <- lm(Bad_moves ~ treat, data = d)
    anova_m1 <- anova(model_1)
    
    return(anova_m1$`Pr(>F)`)
    }

```

## Simulate
```{r Simulate, include = TRUE, message=FALSE}
# Sample sizes to simulate
steps <- seq(0, 400, by=25)

# Effect sizes to simulate
effects <- c(.05, .075, .10, .15)

# Create master data frame to aggregate the data into
g_total <- data.frame()

for (eff in effects){ #loop through effect sizes
  power_list <- data.frame()
  
  # Print for status
  # print(paste(toString(eff*100),"%", sep = ""))
  
  for (step in steps){ #loop through sample sizes 
    
    # simulate_regression <- function(recruits, mean_bad_moves, effect_size, sd_bad_moves)
    p_vals <- replicate(250, simulate_regression(step, 69, eff, 13.7))
    
    power <- length(p_vals[p_vals < 0.05])/length(p_vals)
    power_list <- rbind(power_list, power)
  }
  g <- bind_cols(power_list, steps)
  colnames(g) <- c("y", "x")
 
  # Create column with effect size for this loop 
  x<-rep(c(paste(toString(eff*100),"%", sep = "")),times=length(steps))
  g["Effect_size"] = x
  
  # append data to master data frame
  g_total <- rbind(g_total, g)
}
```
## Plot
```{r Plot, include = TRUE}
g_total$Effect_size <- factor(g_total$Effect_size, levels = c("5%", "7.5%", "10%", "15%"))

plot <- ggplot(data = g_total, aes(x=x, y=y, color = Effect_size)) + 
  geom_line() + 
  geom_point() +
  scale_y_continuous(name="Power", limits=c(0, 1), labels=c("0","20%","40%", "60%","80%","100%"), breaks=seq(0,1,.20)) +
  scale_x_continuous(name="Samples", limits=c(0, 400), labels=c("0","50","100", "150","200","250","300","350", "400"), breaks=seq(0,400,50)) +
  ylab("Power") + 
  geom_hline(yintercept = 0.8, color = "red", linetype="dashed") +
  ggtitle("Sample size vs Power, mean Accuracy = 69%, SD accuracy = 13.8") +
  scale_color_manual(name="Effect Size", values =c("slategrey", "cyan", "royalblue", "purple"))
plot
```


